---
title: "BarPlots, Histograms, and BoxPlots"
format: html
editor: visual
---
 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r include=FALSE}
library(tidyverse)

```


## Barplots

### Exercise 1A
Types of cars and quantity in dataset.
```{r}
ggplot(aes(x = class), data = mpg) +
  geom_bar(aes(fill = drv), position = "dodge")
```

### Exercise 1B
Different types of drives in different types of cars .
*(4-wheel drive, 3-wheel drive, etc.)*
```{r}
ggplot(aes(x = drv), data = mpg) +
  geom_bar(aes(fill = class), position = "dodge")
```

## Histograms

### Exercise 2A
Quantity & price of diamonds.
```{r}
ggplot(aes(x = price), data = diamonds) +
  geom_histogram(binwidth = 250, color = "blue")
```

### Exercise 2B
Density bar graph of price of diamonds. The higher the peak, the more diamonds
of that type there are.
```{r}
ggplot(aes(x = price), data = diamonds) +
  geom_density(color = "blue")
```

### Exercise 2C
Prices of diamonds by color.
```{r}
ggplot(aes(x = price), data = diamonds) +
  geom_histogram(binwidth = 500, color = "blue") + 
  facet_wrap(~color)
```

## BoxPlots

## Exercise 3
Average milleage for different types of cars.
```{r}
mpg %>%
ggplot(aes(x = hwy, y = class)) + 
  geom_boxplot(color = "black")
  
```

